Something something progress in light

### 2022-08-06 2_SSL_EES/article/article_main.tex
added introduction and previous literature

A rapid reduction of global carbon dioxide emissions is urgently required in order to mitigate the effects of climate change \cite{Forster2019}. According to the United Nations, by the end of 2021, more than 130 countries have set or are considering setting a net zero emissions target by or around 2050 \cite{un2021climate}. The European Union, for example, has set a goal of net zero emissions by 2050—a goal it aims to meet with the help of the Green New Deal along other EU and national policies as part of the \textit{European Green Deal} \cite{eu2020green}.

Achieving these ambitious and critically important targets will require both the deployment of new clean energy technologies, and the acceleration of innovation in existing supply-side \cite{sinn2012green} and demand-side technologies \cite{rgeVorsatz2009}. To ensure rapid adoption of these technologies, improvements in their costs, performance, and consumer experience are needed. This requires understanding how cost reductions and performance improvements can be achieved in these technologies \cite{Stephan2021}\cite{kolesnikov2020novel} \cite{Ziegler2021}.

There are a range of mechanisms that contribute to technological change in a technology over time, including targeted research and development (R\&D), economies of scale, learning by doing \cite{Arrow1971}, and economies of scope \cite{johansson2012global}\cite{national2016power}\cite{iea2020perspectives}. The role of these factors at different stages of the innovation life cycle \cite{grubler2012policies}, from research and technology development to demonstration, market formation and diffusion, is an area of active research \cite{Mowery1979} \cite{kavlak2018evaluating} \cite{kolesnikov2020novel} \cite{Ziegler2021}. Innovation is also driven by forces of supply and demand. Various \textit{"technology-push"} drivers reduce the costs of innovating, while \textit{"market-pull"} drivers increase the pay-offs from investing in innovation \cite{anadon2009policy}.

The stages and drivers of the innovation life-cycle for a particular technology play out within a broader innovation system \cite{grubler2012policies} \cite{Anadon2016}. Among these drivers, the role of knowledge transfer and technology spillovers in research and development of energy technologies is an understudied area \cite{Stephan2021}. While the exact definition of spillovers in literature depends on the context and the research question \cite{Liu2003}\cite{Nemet2012} we follow Stephan et al. \cite{Stephan2021} and consider knowledge to be external – thus a spillover –  if it has been developed for application in other technologies, sectors, or scientific disciplines. There is emerging evidence that understanding spillovers and the knowledge network outside a technology may be a fruitful approach to understand \cite{Pichler2020} and shape  \cite{Clark2016} \cite{ Stephan2021} \cite{Sun2021} the future evolution of technologies.

Among demand-side technologies, the provision of lighting is a particularly important area  for climate change mitigation efforts, as it currently accounts for $15-19\%$ of global electricity consumption \cite{Zissis2016}\cite{doe_electricity}. It is also an area of rapid recent technological change: since the introduction of the first commercial white light-emitting diodes (LEDs) in 1996, lighting technology has experienced dramatic efficiency improvements. As shown in Figure 1, thanks to the introduction of LED-based solid-state lighting (SSL), the efficacy of lighting sources has increased by three orders of magnitude in just over 20 years, which is significantly faster than the historical progress observed in previous lighting technologies \cite{weinold2021quantifying}. For comparison, the highest performing light-emitting devices today reach efficacies of 220 lm/W \cite{lumistrips2021mid}, while an incandescent light bulb can only reach efficacies of up to 18 lm/W. Moreover, this rapid improvement in efficiency has been accompanied by a similarly impressive decrease in LED manufacturing costs and retail prices. During the same period, LED retail prices have fallen by two orders of magnitude, as shown in Figure 2.

@@

%%%REFERENCES%%%
\bibliography{auxiliary/rsc_style_files/bibliography_rsc}
\bibliographystyle{bibliography} %the RSC's .bst file

>>>>

A rapid reduction of global carbon dioxide emissions is urgently required in order to mitigate the effects of climate change \cite{Forster2019}. According to the United Nations, by the end of 2021, more than 130 countries have set or are considering setting a net zero emissions target by or around 2050 \cite{un2021climate}. The European Union, for example, has set a goal of net zero emissions by 2050—a goal it aims to meet with the help of the European Green Deal, alongside other EU and national policies \cite{eu2020green}.

Achieving these ambitious and critically important targets will require both the deployment of new clean energy technologies, and the acceleration of innovation in existing supply-side \cite{sinn2012green} and demand-side energy technologies \cite{rgeVorsatz2009}. To ensure rapid adoption of these technologies, significant reductions in their costs and improvements in their performance and consumer experience are needed. This requires understanding how these cost reductions and performance improvements can be achieved \cite{Stephan2021}\cite{Ziegler2021}.

There are a range of mechanisms that contribute to improvements in a technology over time, including targeted research and development (R\&D) efforts, economies of scale, learning by doing \cite{Arrow1971}, and economies of scope \cite{johansson2012global}\cite{national2016power}\cite{iea2020perspectives}. The role of these factors at different stages of the innovation life cycle \cite{grubler2012policies}, from research and technology development to demonstration, market formation and diffusion, is an area of active research \cite{Mowery1979} \cite{kavlak2018evaluating} \cite{Ziegler2021}. Innovation is also driven by forces of supply and demand. Various \textit{"technology-push"} drivers reduce the costs of innovating, while \textit{"market-pull"} drivers increase the pay-offs from investing in innovation \cite{anadon2009policy}.

\begin{figure*}[h]
 \centering
 \includegraphics[height=7cm]{2_SSL_EES/article/figures/history_efficacy.pdf}
 \caption{Historical development of the luminous efficacy ($K$) of the most widely-used lighting technologies in human history from 1850 to 2020. Data shown for best performers by year of market introduction, with dashed lines for visual aid. The physical limit for an ideal light source with a colour rendering index of CRI=90, denoted as $K_{max}^{CRI90}$, is shown as a black horizontal line, as per calculations by Murphy et al. \cite{Murphy2012}.  The magnified plot shows the progress in cool white LEDs from 1996 to 2020, with the dashed line indicating a linear rate of efficacy improvement of 10lm/W per year. For comparison, efficacies of best performers in legacy lighting technologies for 2020 are shown as coloured horizontal lines. Note the logarithmic scale of the vertical axis on the main plot and the linear scale on the magnified plot. Abbreviations: HID – High-Intensity Discharge; CFL – Compact Fluorescent Lamp; Hal. – Halogen, Incd. – Incandescent. Source: own synthesis of published data based on a visual approach proposed by Azevedo et al. \cite{azevedo2009transition}. See Supplementary Information, section X, for the full list of data sources. }
 \label{fgr:history_efficecy}
\end{figure*}

The stages and drivers of the innovation life-cycle for a particular technology play out within a broader innovation system \cite{grubler2012policies} \cite{Anadon2016}. Among these drivers, the role of external knowledge transfer and technology spillovers in research and development of energy technologies is an understudied area \cite{Stephan2021}. While the exact definition of spillovers in literature depends on the context and the research question \cite{liu2003does}\cite{Nemet2012}, we follow Stephan et al. \cite{Stephan2021} and consider knowledge to be external – thus a spillover –  if it has been developed for application in other technologies, sectors, or scientific disciplines. There is emerging evidence that understanding spillovers and the knowledge network outside a technology may be an important factor in understanding \cite{Pichler2020} and shaping \cite{Clark2016} \cite{ Stephan2021} \cite{Sun2021} the future evolution of technologies.

Among demand-side technologies, the provision of lighting is a particularly important area for climate change mitigation efforts, as it currently accounts for $15-19\%$ of global electricity consumption \cite{Zissis2016}\cite{doe_electricity}. It is also an area of rapid recent technological change: since the introduction of the first commercial white light-emitting diodes (LEDs) in 1996, lighting technology has experienced dramatic efficiency improvements. As shown in Figure \ref{fgr:history_efficecy}, thanks to the introduction of LED-based solid-state lighting (SSL), the efficacy of lighting sources has increased by three orders of magnitude in just over 20 years, which is significantly faster than the historical progress observed in previous lighting technologies \cite{weinold2021quantifying}. For comparison, the highest performing light-emitting devices today reach efficacies of 220 lm/W \cite{lumistrips2021mid}, while an incandescent light bulb can only reach efficacies of up to 18 lm/W. Moreover, this rapid improvement in efficiency has been accompanied by a similarly impressive decrease in LED manufacturing costs and retail prices. During the same period, LED retail prices have fallen by two orders of magnitude, as shown in Figure \ref{fgr:cost_lamp_small}.

\begin{figure}
\centering
  \includegraphics[height=5cm]{2_SSL_EES/article/figures/cost_lamp_small.pdf}
  \caption{Historical development of retail sales prices (in 2020 USD) per luminous flux of LED-based luminaires, including light bulbs, spotlights and recessed lights from 2008 to 2020. Annual price per flux decline for the represented period is $27.3\%$, in line with previous estimates \cite{Gerke2020}. Red curved and dashed lines represent average retail sales prices and price projections for LED based luminaires published by the U.S. Department of Energy (DoE) \cite{ council2013assessment}. Shown for reference are the average prices for compact fluorescent (CFL) and incandescent light bulbs. Source: own synthesis of data on LED sales prices collected from various consumer watchdog databases and publications. Data on CFL sales prices collected from \cite{eger2018origin} as well as various consumer watchdog databases and publications. Incandescent light bulb price is assumed constant based on the average in the covered time period. See Supplementary information, section XX, for the full list of data sources.}
  \label{fgr:cost_lamp_small}
\end{figure}

LED sales prices collected from various consumer watchdog databases and publications. Data on CFL sales prices collected from \cite{eger2018origin} as well as various consumer watchdog databases and publications. Incandescent light bulb price is assumed constant based on the average in the covered time period. See Supplementary information, section …, for the full list of data sources.

These dramatic improvements in lighting technology, supported by the introduction of lighting efficiency regulations phasing out incandescent lightbulbs and targeted policies stimulating LED adoption in many countries, led to the rapid expansion and diffusion of SSL technologies \cite{weinold2020long}\cite{Mills2014}\cite{stegmaier2021incandescent}\cite{grubb2021new}. As a result, by 2020, highly efficient LED luminaires were saving an estimated 131 TWh/year for the EU \cite{eu2019impactass} and 442 TWh/year for the US \cite{guidehouse2020adoption}, which is on par with the amount of energy produced annually by all solar photovoltaic installations in these regions. Notably, market adoption of LED lighting is not limited to developed economies \cite{Kamat2020}. For example, durability, low up-front cost and high efficiency of LED light sources have led to their wide-spread adoption in rural West African communities without access to grid electricity \cite{Bensch2017}. LEDs have also been used in a wide range of applications beyond lighting, such as personal health monitors \cite{o2019optical}, \cite{Wyatt2020}, watches and smartphones \cite{Bai2017}, potable water treatment \cite{Lui2014}, high-bandwidth wireless data transmission \cite{Haas2016}, and augmented reality eye wear \cite{Lee2016}.

Despite this impressive history, the sources of LED innovation have not received as much attention from researchers as innovation in supply-side energy technologies, such as solar photovoltaics \cite{kavlak2018evaluating} or wind energy \cite{qiu2012price}\cite{jennings2020policy}, or in lithium-ion batteries for transportation \cite{Ziegler2021}\cite{Stephan2021}. To the best of our knowledge, no study has comprehensively discussed the sources or extent of progress across various metrics of LED cost or performance since the introduction of first commercial white LED products. Understanding the extent to which individual innovations and knowledge spillovers contributed to improvements in LED technology, how this effect compares to other sources of improvements such as economies of scale, and how these innovations occurred (i.e., by what mechanisms and actors) will provide valuable lessons both for other demand-side technologies and overall for accelerating clean energy innovation.

To address these questions, in this paper we identify a set of metrics suitable for tracking the historical progress of LED lighting technology. We then trace device efficiency and cost improvements, as well as changes in relevant consumer experience metrics from the time of introduction of the first commercial warm white LEDs in 2003 to 2020, the year with the most recent data available at the time of writing. Given the proprietary nature of knowledge in the SSL industry, we collect corresponding information using a multi-method approach combining a systematic literature review of scientific literature and industry reports with patent analysis and a series of elite interviews \cite{tansey2009process} with eminent experts from academia, public research institutions and industry. This approach allows us to identify innovations in white LED technology and examine their origins to discern technology spillovers among them. Data on performance improvements associated with individual innovations identified in patents, scientific publications, industry reports, and interviews is then augmented by our calculations of device efficiency, including contributions of specific spillovers to the overall LED efficiency (see Supplementary Information for more detail). We further calculate LED manufacturing costs using own bottom-up cost model with a process-step resolution. Finally, we discuss what our observations and conclusions mean for the understanding of the innovation process in LED technology and for broader policy and industry efforts to accelerate clean energy innovation.

\section{Previous literature}

The historical development of light-emitting diodes has received a considerable amount of attention following the recognition of pioneering work on blue LED by Japanese researchers Shuji Nakamura, Isamu Akasaki and Hiroshi Amano with the 2014 Nobel Prize in Physics \cite{Akasaki2015} \cite{Nakamura2015}. The sources and disaggregated contributions of subsequent improvements in LED technology, however, have not been documented in the literature in a systematic fashion, though there were notable publications reporting on the progress in the design of devices \cite{Shchekin2006} \cite{krames2007led} \cite{laubsch2009high} \cite{hahn2014development}, or contributions of different sub-efficiencies to overall device efficiency for 2009 \cite{tsao2010solid} and 2016 \cite{pattison2017solid}.

Several descriptive papers include very useful information on some of the scientific breakthroughs that have contributed to LED progress \cite{krames2007status}, \cite{Phillips2007}, \cite{Bierhuizen2007}, \cite{Nakamura2013}, \cite{feezell2018invention} and \cite{Taki2019}. Additional  studies providing more detail on specific advances in LED technology were also published by authors working for key industry actors Lumileds\cite{MuellerMach2005}, \cite{Shchekin2006}, \cite{lumi2015lumi}, \cite{Bhardwaj2017} and Osram (now amsOsram) \cite{Haerle2004}, \cite{Baur2009}, \cite{laubsch2009high}, \cite{hahn2014development}. However, these studies typically include limited information on the macro chip-design and cover disparate aspects of the technology over different time periods. Another important source of historical information comes from patents, which cover the entirety of the device architecture or manufacturing process, including those by Lumileds \cite{margalith2011thin}, Samsung \cite{jung2014phos} \cite{cha2019semiconductor} and Soraa \cite{cich2017high}.

We summarize the information collected, analysed and classified on the progress in LED chip architectures and manufacturing processes from these and other disparate sources in Figure 3, where we show the evolution of LEDs from classical chips with lateral current spreading to chip-scale package flip-chip architectures. Despite the amount of literature published on the topic of LED history, to the best of our knowledge, no publication comprehensively aggregates known chip design, manufacturing, and material improvements to show the overall effect of these improvements on device efficiency or manufacturing cost over time. In addition, the effect of individual innovations and technology spillovers, which has been investigated in solar PV \cite{kavlak2018evaluating} \cite{kolesnikov2020novel} \cite{nemet2019solar}, and to some extent in lithium-ion batteries \cite{Stephan2021}, has not been studied in the context of lighting yet. This is consistent with previous observations regarding the marginalization of end-use technologies in the analysis of energy innovation for climate change impact mitigation \cite{Wilson2012}, \cite{Creutzig2018}.

@@

%%%REFERENCES%%%
\bibliography{bibliography}
\bibliographystyle{auxiliary/rsc_style_files/bibliography_rsc} %the RSC's .bst file
